\large How resource abundance and stochasticity affect animals' space-use requirements

### 2024-03-28 writing/appendix-B-simulations.Rmd
updated appendices and methods for empirical model

# attach all necessary packages
library('ctmm')    # for generating movement models and movement modeling
library('terra')   # for working with raster data
library('dplyr')   # for data wrangling
library('purrr')   # for functional programming
library('tidyr')   # for data wrangling (e.g., nested tibbles)
library('mgcv')    # for empirical Bayes GAMs
library('ggplot2') # for fancy plots
library('cowplot') # for fancy multi-panel plots
library('dagitty') # for directed acyclical graphs
library('ggdag')   # for directed acyclical graphs
theme_set(theme_bw()) # change default theme

@@

```{r, eval=TRUE, echo=FALSE}
# attach all necessary packages
library('ctmm')    # for generating movement models and movement modeling
library('terra')   # for working with raster data
library('dplyr')   # for data wrangling
library('purrr')   # for functional programming
library('tidyr')   # for data wrangling (e.g., nested tibbles)
library('mgcv')    # for empirical Bayes GAMs
library('ggplot2') # for fancy plots
library('cowplot') # for fancy multi-panel plots
library('dagitty') # for directed acyclical graphs
library('ggdag')   # for directed acyclical graphs
theme_set(theme_bw()) # change default theme

@@

```{r dag, eval=TRUE, echo = FALSE, fig.cap="Directed Acyclical Graph assumed for inferring the causal effects of $\\e(R)$ and $\\var(R)$ on $H$.", fig.height=4, fig.width=4}
dag <- dagitty('dag{
"E(R)" -> R <- "Var(R)"
"E(R)" -> "Var(R)"
R -> H
}')
coordinates(dag) <- list(x = c(R = 0, 'E(R)' = -1, 'Var(R)' = 1, H = 0),
                         y = c(R = 1, 'E(R)' = 0, 'Var(R)' = 0, H = 2))
ggdag(dag) +
  theme_dag()
```

```{r, eval=TRUE, fig.height=8, fig.width=8, out.width='65%'}
# marginal effects of mu and sigma only (linear on the log link scale)
m_1 <- gam(
  list(
    # predictor for the mean
    hr ~ quantile + mu + sigma2,
    # predictor for the scale (variance = scale * mean)
    ~ quantile + mu + sigma2),
  family = gammals(),
  data = sims,
  method = 'REML')

@@

# ti term of the scale parameter looks over-fit and too uncertain
plot(m_4, all.terms = TRUE, pages = 1, scheme = 3, scale = 0)
```

@@

```{r gamls-terms, fig.cap = "Marginal and interaction effects of each of the terms present in the final GAMLS (on the log link scales). The dashed lines indicate the 95\\% Bayesian credible intervals. The data points are included as rug plots in all marginal terms and as points in the interaction terms.", eval=TRUE, fig.height=16, fig.width=8, out.width='65%', echo=FALSE}
# using custom layout to make figure easier to read
layout(matrix(c(7, 1, 2, 3, 8, 4, 5, 6), nrow = 4, byrow = TRUE))
plot(m, all.terms = TRUE, scheme = 3, scale = 0)
```

>>>>

# attach all necessary packages
library('ctmm')    # for generating movement models and movement modeling
library('terra')   # for working with raster data
library('dplyr')   # for data wrangling
library('purrr')   # for functional programming
library('tidyr')   # for data wrangling (e.g., nested tibbles)
library('mgcv')    # for empirical Bayes GAMs
library('ggplot2') # for fancy plots
library('cowplot') # for fancy multi-panel plots
library('dagitty') # for directed acyclical graphs
library('ggdag')   # for directed acyclical graphs
library('gratia')  # for ggplot-based GAM figures
theme_set(theme_bw()) # change default theme

@@

```{r, eval=TRUE, echo=FALSE}
# attach all necessary packages
library('ctmm')    # for generating movement models and movement modeling
library('terra')   # for working with raster data
library('dplyr')   # for data wrangling
library('purrr')   # for functional programming
library('tidyr')   # for data wrangling (e.g., nested tibbles)
library('mgcv')    # for empirical Bayes GAMs
library('ggplot2') # for fancy plots
library('cowplot') # for fancy multi-panel plots
library('dagitty') # for directed acyclical graphs
library('ggdag')   # for directed acyclical graphs
library('gratia')  # for ggplot-based GAM figures
theme_set(theme_bw()) # change default theme

@@

```{r dag, eval=TRUE, echo = FALSE, fig.cap="Directed Acyclical Graph assumed for inferring the causal effects of $\\e(R)$ and $\\var(R)$ on $H$.", fig.height=4, fig.width=6, out.width='50%'}
dag <- dagitty('dag{
"E(R)" -> R <- "Var(R)"
"E(R)" -> "Var(R)"
R -> H
}')
coordinates(dag) <- list(x = c(R = 0, 'E(R)' = -1, 'Var(R)' = 1, H = 0),
                         y = c(R = 1, 'E(R)' = 0, 'Var(R)' = 0, H = 2))
ggdag(dag) +
  theme_dag()
```

```{r, eval=TRUE, fig.height=7.5, fig.width=12, out.width='\\linewidth'}
# marginal effects of mu and sigma only (linear on the log link scale)
m_1 <- gam(
  list(
    # predictor for the mean
    hr ~ quantile + mu + sigma2,
    # predictor for the scale (variance = scale * mean)
    ~ quantile + mu + sigma2),
  family = gammals(),
  data = sims,
  method = 'REML')

@@

# ti term of the scale parameter looks over-fit and too uncertain
draw(m_4) & theme_bw() + theme(panel.grid = element_blank())
```

@@

```{r gamls-terms, fig.cap = "Marginal and interaction effects of each of the terms present in the final GAMLS (on the log link scales). The dashed lines indicate the 95\\% Bayesian credible intervals. The data points are included as rug plots in all marginal terms and as points in the interaction terms.", eval=TRUE, fig.height=7.5, fig.width=12, out.width='\\linewidth', echo=FALSE}
# using custom layout to make figure easier to read
layout(matrix(c(7, 1, 2, 3, 8, 4, 5, 6), nrow = 4, byrow = TRUE))
draw(m) & theme_bw() + theme(panel.grid = element_blank())
```
